---
title: "IBM_Valuation.RMD"
author: "Fahad Taimur"
date: "3/28/2020"
output: 
  html_document:
    df_print: paged
    theme: flatly
    highlight: tango
    toc: yes
    toc_depth: 2
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  message= FALSE, 
  warning= FALSE,
  out.width = "100%",
  out.height = "100%",
  dpi = 300)
```

# Load the libraries / PDF variable
```{r}
library(rJava)
library(tabulizer)
library(tidyverse)
library(writexl)
library(plotly)
library(readxl)

# Useful resources
# https://www.ibm.com/annualreport/assets/downloads/IBM_Annual_Report_2019.pdf
# https://cran.r-project.org/web/packages/tabulizer/vignettes/tabulizer.html

pdf <- "https://www.ibm.com/annualreport/assets/downloads/IBM_Annual_Report_2019.pdf"
```

# 1.0 Financial Highlights - Page 25
```{r}
# First half of the financial highlights table
financial_highlights1 <- extract_areas(pdf, pages = 25, output = "data.frame")
financial_highlights1

# Second half
financial_highlights2 <- extract_areas(pdf, pages = 25, output = "data.frame")
financial_highlights2

# Data Cleaning and Binding
financial_highlights1 %>%
  as.data.frame() %>%
  filter(For.the.year.ended.December.31. != "Earnings per share of common stock—continuing operations") %>%
  select(-X, -X.1) %>%
  rename(
    Concept = For.the.year.ended.December.31.,
    FY2019 = X2019,
    FY2018 = X2018
  ) %>%
  bind_rows(
    financial_highlights2 %>%
      as.data.frame() %>%
      # Some of the 2018 and 2019 data is in the X and X.1 column
      mutate(X2019 = case_when(
        X2019 == "" ~ X,
        TRUE ~ X2019
      )) %>%
      mutate(X2018 = case_when(
        X2018 == "" ~ X.1,
        TRUE ~ X2018
      )) %>%
      select(-X, -X.1) %>%
      rename(
    Concept = At.December.31.,
    FY2019 = X2019,
    FY2018 = X2018
      )
  ) -> financial_highlights

# Removing special characters and converting to numeric
financial_highlights %>%
  mutate(
    # Remove special characters
    FY2019 = str_remove_all(FY2019, pattern = "\\,|\\$|\\*") %>% as.numeric(),
    FY2018 = str_remove_all(FY2018, pattern = "\\,|\\$|\\*") %>% as.numeric()
  ) %>%
  filter(Concept != "") -> financial_highlights_clean

financial_highlights_clean %>% writexl::write_xlsx(path = "Data_Extracts/FinancialHighlights.xlsx")
financial_highlights_clean 
```

# 2.0 Revenue / GM by Segment ($ in Millions) - Page 35
```{r}
segment_revenue_gm <- extract_areas(pdf, pages = 35, output = "data.frame")

# Header Names
headers <- c("Segment", "FY2019", "FY2018", "Change", "Adj_for_currency")

# Subset
segment_revenue_gm %>%
  as.data.frame() %>%
  slice(-4:length(segment_revenue_gm$X)) -> segment_revenue_gm

# Change column names
colnames(segment_revenue_gm) <- headers

# Clean FY2018 and 2019
segment_revenue_gm %>%
  mutate(
    FY2019 = str_remove_all(FY2019, pattern = "\\,|\\$|\\*|\\%"),
    FY2018 = str_remove_all(FY2018, pattern = "\\,|\\$|\\*|\\%")
  ) -> segment_revenue_gm_clean

# Let's separate gross margins
segment_revenue_gm_clean %>%
  mutate(Segment_Lag = lag(Segment)) %>%
  filter(Segment == "Gross margin") %>%
  select(Concept = Segment, Segment = Segment_Lag,-Change, -Adj_for_currency, FY2019, FY2018) -> segment_gm_clean

# Gross Margin by Segment
segment_gm_clean
segment_gm_clean %>%
  writexl::write_xlsx("Data_Extracts/BusinessSegmentGrossMarginPercent.xlsx")

# Revenue by Segment
segment_revenue_gm_clean %>%
  filter(Segment != "Gross margin") -> segment_revenue_clean

segment_revenue_clean 

segment_revenue_clean %>%
  writexl::write_xlsx("Data_Extracts/BusinessSegmentRevenue.xlsx")
```

# 2.1 Segment Charts - Revenue down by ~ $ 3M 
```{r fig.height= 8}
segment_revenue_clean %>%
  select(Segment:FY2018) %>%
  gather(key = Year, value = "value", -Segment) %>%
  mutate(value = as.numeric(value)) %>%
  arrange(Segment) %>%
  mutate(year = case_when(
    Year == "FY2019" ~ 2019,
    TRUE ~ 2018
  )) %>%
  mutate(Segment = Segment %>% as_factor() %>% fct_reorder(-value)) %>%
  ggplot(aes(x = factor(year), y = value, fill = Segment )) + 
  geom_col(show.legend = F) + 
  facet_wrap(~ Segment) +
  theme_light() +
  scale_y_continuous(labels = scales::comma_format()) + 
  theme(
     # rectangle background of the facets
    strip.background = element_rect(color = "black", fill = "black", size = 1),
    # working on the text in the strips
    strip.text = element_text(face = "italic")
  ) + 
  labs(title = "Revenue by Segment and Total - IBM", 
       caption = "Global Technology and Cloud/Cognitive  \n are the biggest segments") +
  geom_text(aes(label = scales::dollar(value, scale = 1e-3, suffix = "M", accuracy = 1)), vjust = 1.5, color = "white")-> g

g
```

# 2.2 Save as PNG
```{r}
png(file="Images/segment_revenues.png", width=800, height=500)
g
dev.off()
```

# 3.0 Geographic Revenues
```{r}
areas <- locate_areas(pdf, pages = 40)

geo_revenues <- extract_tables(pdf, 
                               pages = 40, 
                               area = list(c(291, 88, 345, 512)),
                               guess = F, 
                               output = "data.frame") %>%
                as.data.frame()



geo_revenues %>%
  as.data.frame() %>%
  rename(
    Region = names(geo_revenues[1]),
    FY2019 = names(geo_revenues[2]),
    FY2018 = names(geo_revenues[3])
  ) %>%
    mutate(
    FY2019 = str_remove_all(FY2019, pattern = "\\,|\\$|\\*|\\%") %>% as.numeric(),
    FY2018 = str_remove_all(FY2018, pattern = "\\,|\\$|\\*|\\%") %>% as.numeric()
  ) -> geo_revenues_clean

geo_revenues_clean

geo_revenues_clean %>%
  writexl::write_xlsx("Data_Extracts/RevenueByGeography.xlsx")
```

# 3.1 Geographic Chart of Revenues
```{r}
geo_revenues_clean %>%
  select(Region:FY2018) %>%
  gather(key = Year, value = "value", -Region) %>%
  mutate(year = case_when(
    Year == "FY2019" ~ 2019,
    TRUE ~ 2018
  )) %>%
  mutate(Region = Region %>% as_factor() %>% fct_reorder(-value)) %>%
  ggplot(aes(x = factor(year), y = value, fill = Region )) + 
  geom_col(show.legend = F) + 
  facet_wrap(~ Region) +
  theme_light() +
  scale_y_continuous(labels = scales::comma_format()) + 
  theme(
     # rectangle background of the facets
    strip.background = element_rect(color = "black", fill = "black", size = 1),
    # working on the text in the strips
    strip.text = element_text(face = "italic")
  ) + 
  labs(title = "Revenue by Geography", 
       caption = "Americas represents ~ 50% of IBM Revenue") +
  geom_text(aes(label = scales::dollar(value, scale = 1e-3, suffix = "M", accuracy = 1)), vjust = 1.5, color = "white")-> g

g
```

# 3.2 Save as PNG 
```{r}
png(file="Images/RevenueByGeography.png", width=800, height=500)
g
dev.off()
```


# 3.3 Five Year Comparison of Select Financial Data (Page 140)
```{r}
areas <- locate_areas(pdf, pages = 140)

five_year_comp <- extract_tables(pdf, 
                               pages = 140, 
                               area = list(c(79, 88, 301, 512)),
                               guess = F, 
                               output = "data.frame") %>%
                as.data.frame()

# Wide format
five_year_comp %>%
  select(-X, -X.1, -X.2, -X.3, -X.4) %>%
  filter(X2019. != "") %>%
  write.csv(file = "Data_Extracts/five_year_comparison_of_key_financials_pg140.csv")
# Tidy format


five_year_comp %>%
  select(-X, -X.1, -X.2, -X.3, -X.4) %>%
  filter(X2019. != "") %>%
  gather(key = Year, value = Value, -names(five_year_comp[1])) %>%
  rename(Concept = "For.the.year.ended.December.31.") %>%
  arrange(Concept) %>%
  mutate(Year = str_remove_all(Year, pattern = "X") %>% as.numeric())

```

# 4.0 Business Units
# 4.1 Cloud / Cognitive Software
```{r}
# Locate the areas and vectorize them
areas <- locate_areas(pdf, pages = 35)
areas <- areas %>%
  pluck(1) %>%
  as.vector()

areas

# Get the data
cloud_cognitive <- extract_tables(pdf, 
                               pages = 35, 
                               area = list(areas),
                               guess = F, 
                               output = "data.frame") %>%
                as.data.frame()

# Clean
cloud_cognitive %>%
    rename(
    SubArea = names(cloud_cognitive[1]),
    FY2019 = names(cloud_cognitive[2]),
    FY2018 = names(cloud_cognitive[3])
  ) %>%
    mutate(
    FY2019 = str_remove_all(FY2019, pattern = "\\,|\\$|\\*|\\%") %>% as.numeric(),
    FY2018 = str_remove_all(FY2018, pattern = "\\,|\\$|\\*|\\%") %>% as.numeric()
  ) -> cloud_cognitive_clean; cloud_cognitive_clean
  

# Save
cloud_cognitive_clean %>%
  write.csv("Data_Extracts/cloud_cognitive_softwares.csv")

# Plot
cloud_cognitive_clean %>%
  select(SubArea:FY2018) %>%
  gather(key = Year, value = "value", -SubArea) %>%
  mutate(year = case_when(
    Year == "FY2019" ~ 2019,
    TRUE ~ 2018
  )) %>%
  mutate(SubArea = SubArea %>% as_factor() %>% fct_reorder(-value)) %>%
  ggplot(aes(x = factor(year), y = value, fill = SubArea )) + 
  geom_col(show.legend = F) + 
  facet_wrap(~ SubArea) +
  theme_light() +
  scale_y_continuous(labels = scales::comma_format()) + 
  theme(
     # rectangle background of the facets
    strip.background = element_rect(color = "black", fill = "black", size = 1),
    # working on the text in the strips
    strip.text = element_text(face = "italic")
  ) + 
  labs(title = "Revenue by Cloud & Cognitive Segment", 
       caption = "Cloud / Cognitive Segment with a billion dollar revenue increase") +
  geom_text(aes(label = scales::dollar(value, scale = 1e-3, suffix = "B", accuracy = 0.1)), vjust = 1.5, color = "white")-> g

# Preview
g

# Save Image
png(file="Images/CloudCognitiveRevenue.png", width=800, height=500)
g
dev.off()
```

# 4.2 Global Technology Services (GTS)
```{r}
# Locate the areas and vectorize them
areas <- locate_areas(pdf, pages = 37)
areas <- areas %>%
  pluck(1) %>%
  as.vector()

areas

# Get the data
gts <- extract_tables(pdf, 
                               pages = 37, 
                               area = list(areas),
                               guess = F, 
                               output = "data.frame") %>%
                as.data.frame()

# Clean
gts %>%
    rename(
    SubArea = names(gts[1]),
    FY2019 = names(gts[2]),
    FY2018 = names(gts[3])
  ) %>%
    mutate(
    FY2019 = str_remove_all(FY2019, pattern = "\\,|\\$|\\*|\\%") %>% as.numeric(),
    FY2018 = str_remove_all(FY2018, pattern = "\\,|\\$|\\*|\\%") %>% as.numeric()
  ) -> gts_clean; gts_clean
  

# Save
gts_clean %>%
  write.csv("Data_Extracts/global_technology_services.csv")

# Plot
gts_clean %>%
  select(SubArea:FY2018) %>%
  gather(key = Year, value = "value", -SubArea) %>%
  mutate(year = case_when(
    Year == "FY2019" ~ 2019,
    TRUE ~ 2018
  )) %>%
  mutate(SubArea = SubArea %>% as_factor() %>% fct_reorder(-value)) %>%
  ggplot(aes(x = factor(year), y = value, fill = SubArea )) + 
  geom_col(show.legend = F) + 
  facet_wrap(~ SubArea) +
  theme_light() +
  scale_y_continuous(labels = scales::comma_format()) + 
  theme(
     # rectangle background of the facets
    strip.background = element_rect(color = "black", fill = "black", size = 1),
    # working on the text in the strips
    strip.text = element_text(face = "italic")
  ) + 
  labs(title = "Revenue by Global Technology Services", 
       caption = "Decline of ~ 1.7 billion in GTS") +
  geom_text(aes(label = scales::dollar(value, scale = 1e-3, suffix = "B", accuracy = 0.1)), vjust = 1.5, color = "white")-> g

# Preview
g

# Save Image
png(file="Images/GlobalTechServices.png", width=800, height=500)
g
dev.off()
```

# 4.3 Systems Revenue 
```{r}
# Locate the areas and vectorize them
areas <- locate_areas(pdf, pages = 38)
areas <- areas %>%
  pluck(1) %>%
  as.vector()

areas

# Get the data
systems <- extract_tables(pdf, 
                         pages = 38, 
                         area = list(areas),
                         guess = F, 
                         output = "data.frame") %>%
                as.data.frame()

# Clean
systems %>%
    rename(
    SubArea = names(systems[1]),
    FY2019 = names(systems[2]),
    FY2018 = names(systems[3])
  ) %>%
    mutate(
    FY2019 = str_remove_all(FY2019, pattern = "\\,|\\$|\\*|\\%") %>% as.numeric(),
    FY2018 = str_remove_all(FY2018, pattern = "\\,|\\$|\\*|\\%") %>% as.numeric()
  ) -> systems_clean; systems_clean
  

# Save
systems_clean %>%
  write.csv("Data_Extracts/SystemsRevenue.csv")

# Plot
systems_clean %>%
  select(SubArea:FY2018) %>%
  na.omit() %>%
  gather(key = Year, value = "value", -SubArea) %>%
  mutate(year = case_when(
    Year == "FY2019" ~ 2019,
    TRUE ~ 2018
  )) %>%
  mutate(SubArea = SubArea %>% as_factor() %>% fct_reorder(-value)) %>%
  ggplot(aes(x = factor(year), y = value, fill = SubArea )) + 
  geom_col(show.legend = F) + 
  facet_wrap(~ SubArea) +
  theme_light() +
  scale_y_continuous(labels = scales::comma_format()) + 
  theme(
     # rectangle background of the facets
    strip.background = element_rect(color = "black", fill = "black", size = 1),
    # working on the text in the strips
    strip.text = element_text(face = "italic")
  ) + 
  labs(title = "Revenue of Systems", 
       caption = "Decline in Systems Hardware segement") +
  geom_text(aes(label = scales::dollar(value, scale = 1e-3, suffix = "B", accuracy = 0.1)), vjust = 1.5, color = "white")-> g

# Preview
g

# Save Image
png(file="Images/SystemsRevenue.png", width=800, height=500)
g
dev.off()
```

# 4.4 Global Busines Services 
```{r}
# Locate the areas and vectorize them
areas <- locate_areas(pdf, pages = 36)
areas <- areas %>%
  pluck(1) %>%
  as.vector()

areas

# Get the data
gbs <- extract_tables(pdf, 
                     pages = 36, 
                     area = list(areas),
                     guess = F, 
                     output = "data.frame") %>%
                as.data.frame()

# Clean
gbs %>%
    rename(
    SubArea = names(gbs[1]),
    FY2019 = names(gbs[2]),
    FY2018 = names(gbs[3])
  ) %>%
    mutate(
    FY2019 = str_remove_all(FY2019, pattern = "\\,|\\$|\\*|\\%") %>% as.numeric(),
    FY2018 = str_remove_all(FY2018, pattern = "\\,|\\$|\\*|\\%") %>% as.numeric()
  ) -> gbs_clean; gbs_clean
  

# Save
gbs_clean %>%
  write.csv("Data_Extracts/global_business_services.csv")

# Plot
gbs_clean %>%
  select(SubArea:FY2018) %>%
  gather(key = Year, value = "value", -SubArea) %>%
  mutate(year = case_when(
    Year == "FY2019" ~ 2019,
    TRUE ~ 2018
  )) %>%
  mutate(SubArea = SubArea %>% as_factor() %>% fct_reorder(-value)) %>%
  ggplot(aes(x = factor(year), y = value, fill = SubArea )) + 
  geom_col(show.legend = F) + 
  facet_wrap(~ SubArea) +
  theme_light() +
  scale_y_continuous(labels = scales::comma_format()) + 
  theme(
     # rectangle background of the facets
    strip.background = element_rect(color = "black", fill = "black", size = 1),
    # working on the text in the strips
    strip.text = element_text(face = "italic")
  ) + 
  labs(title = "Revenue by Global Business Services", 
       caption = "GBS segment has maintained sales from 2018 to 2019") +
  geom_text(aes(label = scales::dollar(value, scale = 1e-3, suffix = "B", accuracy = 0.1)), vjust = 1.5, color = "white")-> g

# Preview
g

# Save Image
png(file="Images/GlobalBusinessServices.png", width=800, height=500)
g
dev.off()
```

# 5.0 Regression Analysis - IBM Brand Value
# 5.1 Data
```{r}
# Let's use the 5-years of data 
bv_data <- readxl::read_excel("IBM_Brand_Value.xlsx")

bv_data %>%
  filter(Year %>% between(2005, 2013)) -> reg_data

bv_ts <- ts(data = reg_data[,2], start = 2005, frequency = 1)
bv_ts
```

# 5.2 Regression
```{r}
library(tseries)
library(forecast)

# Perform time-series regression of the brand value
reg <- tslm(BrandValue ~ trend, data = bv_ts)
summary(reg)
```
# 5.3 Regression Diagnostics
```{r}
# Setting up a hypothesis test for autocorrelation:
# null: there is no autocorrelation in the residuals
# alternative: there is serial autocorrelation
# There is not significant autocorrelation as shown by the p-value and ACF plot

# Residual distribution is not normal and residuals display a varying pattern

checkresiduals(reg)

# Testing the magnitude of error
# RMSE of ~2.1B$ 
bind_cols(Actual = bv_ts[,1], Fitted = reg$fitted.values) %>%
  as_tibble() %>%
  mutate(Actual = as.numeric(Actual), 
         Fitted = as.numeric(Fitted)) %>%
  yardstick::metrics(truth = Actual, estimate = Fitted)

# % error ~ 3%

rmse <- 2100
rmse/mean(bv_data %>% filter(Year %>% between(2005, 2013)) %>% select(BrandValue) %>% pull())
```

# 5.3 Forecast vs. Actual
```{r}
bv_forecast <- forecast(reg, h = 6)

# Bind the Actual to the Forecasted values
bv_forecast %>% 
  as.tibble() %>% 
  janitor::clean_names() %>%
  select(Forecast = point_forecast) %>%
  pull(Forecast)-> bv_forecast

bv_forecast_vector <- c(rep.int(NA_integer_, 8), 78808, bv_forecast)

bv_data %>%
  filter(Year %>% between(2005, 2019)) -> bv_data

bind_cols(
  Year = bv_data$Year,
  Actual_BV = bv_data$BrandValue,
  Forecasted = round(bv_forecast_vector)
) -> bv_actual_forecasted

bv_actual_forecasted %>%
  write_csv("brand_value_actual_vs_forecasted.csv")
```

# Plot of Demise
```{r}
dygraphs::dygraph(bv_actual_forecasted)
```





